---
title: "Untitled"
author: "Jintong YAN"
date: "2023-04-17"
output: pdf_document
---
```{r setup, include=FALSE}
library(survminer)
library(coda)
library(dplyr)
library(tidyverse)
library(edgeR)
library("readxl")
library(ggplot2)
library(HDInterval)
library(ggpubr)
library(JM)
library("lattice")
theme_set(
  cowplot::theme_cowplot() + 
    theme(
      strip.background = element_blank(), 
      strip.text = element_text(size = 8),
      legend.position = 'bottom'
    )
)
knitr::opts_chunk$set(echo = T, fig.height = 5, fig.width = 7, dpi = 300, dev = 'svglite')
```

# An Example Based on Poisson distribution with a Gamma prior

In this model, the parameter of interest is the rate parameter of a Poisson distribution. The prior distribution is a Gamma distribution with parameters $\alpha$ and $\beta$, and the likelihood function is a Poisson distribution with parameter $\lambda$. The posterior distribution is also a Gamma distribution with updated parameters  $\alpha'$ and $\beta'$ that depend on the sum of the observed counts.

In this example, first, we generate 10 observations from the distribution with a known parameter $\lambda=5$.

```{r}
n <- 5
lambda_true <- 2.2
set.seed(123) 
y <- rpois(n, lambda_true)
y
```

Then, we forget the true value and try to use the prior we choose and the data we have(y) to get estimate the posterior distribution of the parameter from this data. 

```{r}
### prior distribution: Gamma(1, 1)
alpha <- 1
beta <- 1
lambda <- seq(0, 10, by = 0.01) # set up grid for plotting
### prior distribution plot
plot(lambda, dgamma(lambda, alpha, beta), type = 'l', lwd = 2, col = 'orange',
     ylim = c(0, 3.2), xlab = expression(lambda), 
     ylab = expression(paste('p(', lambda, '|y)')))
### posterior distribution plot
lines(lambda, dgamma(lambda, alpha + sum(y), beta + n), 
      type = 'l', lwd = 2, col = 'violet')
abline(v = lambda_true, lty = 2)
legend('topright', inset = .02, legend = c('prior', 'posterior'),
       col = c('orange', 'violet'), lwd = 2)
```

## Equal-tailed interval
### prior distribution: Gamma(1, 1)
95% confidence interval for the parameter $\lambda$
```{r}
alpha <- 1 
beta <- 1
conf <- 0.05
pdf("/Users/yanjintong/2023/STAT520C/gamma1-1.pdf", width=8, height=8)
plot(lambda, dgamma(lambda, alpha, beta), type = 'l', lwd = 2, col = 'orange')
lambda <- seq(0, 15, by = 0.001) # set up grid for plotting
```

n = 2

```{r}
n <- 2
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_1 <- sum(y) + alpha
beta_1 <- n + beta
q_lower_1 <- qgamma(conf / 2, alpha_1, beta_1)
q_upper_1 <- qgamma(1 - conf / 2, alpha_1, beta_1)
c(q_lower_1, q_upper_1)
```

n = 10

```{r}
n <- 5
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_2 <- sum(y) + alpha
beta_2 <- n + beta
q_lower_2 <- qgamma(conf / 2, alpha_2, beta_2)
q_upper_2 <- qgamma(1 - conf / 2, alpha_2, beta_2)
c(q_lower_2, q_upper_2)
```

n = 15

```{r}
n <- 15
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_3 <- sum(y) + alpha
beta_3 <- n + beta
q_lower_3 <- qgamma(conf / 2, alpha_3, beta_3)
q_upper_3 <- qgamma(1 - conf / 2, alpha_3, beta_3)
c(q_lower_3, q_upper_3)
```

n = 75

```{r}
n <- 75
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_4 <- sum(y) + alpha
beta_4 <- n + beta
q_lower_4 <- qgamma(conf / 2, alpha_4, beta_4)
q_upper_4 <- qgamma(1 - conf / 2, alpha_4, beta_4)
c(q_lower_4, q_upper_4)
```

Put them in one plot

```{r}
plot_CI <- function(alpha, beta, conf, lambda_true, n) {
    plot(lambda, dgamma(lambda, alpha, beta), type = 'l', lwd = 2, col = 'violet',
       ylim = c(0, 2), xlim = c(0, 10), xlab = expression(lambda),
    ylab = expression(paste('p(', lambda, '|y)')))
    q_lower <- qgamma(conf / 2, alpha, beta)
    q_upper <- qgamma(1 - conf / 2, alpha, beta)
    y_val <- dgamma(lambda, alpha, beta)
    x_coord <- c(q_lower, lambda[lambda >= q_lower & lambda <= q_upper], q_upper)
    y_coord <- c(0, y_val[lambda >= q_lower & lambda <= q_upper], 0)
    polygon(x_coord, y_coord, col = 'pink', lwd = 2,  border = 'violet')
    lines(lambda, dgamma(lambda, alpha, beta),
        type = 'l', lwd = 2, col = 'orange')
    abline(v = lambda_true, lty = 2)
    text(x = 0.5, y = 1.5, paste0('n=', n), cex = 1.75)
}
pdf("/Users/yanjintong/2023/STAT520C/ETI_1.pdf", width=8, height=8)
par(mfrow = c(2,2), mar = c(2, 2, .1, .1))
plot_CI(alpha_1, beta_1, conf = 0.05, lambda_true, n=2)
plot_CI(alpha_2, beta_2, conf = 0.05, lambda_true, n=5)
plot_CI(alpha_3, beta_3, conf = 0.05, lambda_true, n=15)
plot_CI(alpha_4, beta_4, conf = 0.05, lambda_true, n=75)
```

### prior distribution: Gamma(8, 1)
compute 95% confidence interval for the parameter $\lambda$
```{r}
# Prior distribution 
alpha <- 8
beta <- 1
pdf("/Users/yanjintong/2023/STAT520C/gamma8-1.pdf", width=8, height=8)
plot(lambda, dgamma(lambda, alpha, beta), type = 'l', lwd = 2, col = 'orange')
```

```{r}
# posterior distribution
n <- 2
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_1 <- sum(y) + alpha
beta_1 <- n + beta
q_lower <- qgamma(conf / 2, alpha_1, beta_1)
q_upper <- qgamma(1 - conf / 2, alpha_1, beta_1)
c(q_lower, q_upper)
n <- 5
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_2 <- sum(y) + alpha
beta_2 <- n + beta
q_lower <- qgamma(conf / 2, alpha_2, beta_2)
q_upper <- qgamma(1 - conf / 2, alpha_2, beta_2)
c(q_lower, q_upper)
n <- 15
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_3 <- sum(y) + alpha
beta_3 <- n + beta
q_lower <- qgamma(conf / 2, alpha_3, beta_3)
q_upper <- qgamma(1 - conf / 2, alpha_3, beta_3)
c(q_lower, q_upper)
n <- 75
set.seed(123) 
y <- rpois(n, lambda_true)
# posterior distribution for the parameter lambda
alpha_4 <- sum(y) + alpha
beta_4 <- n + beta
q_lower <- qgamma(conf / 2, alpha_4, beta_4)
q_upper <- qgamma(1 - conf / 2, alpha_4, beta_4)
c(q_lower, q_upper)

pdf("/Users/yanjintong/2023/STAT520C/ETI_2.pdf", width=8, height=8)
par(mfrow = c(2, 2), mar = c(2, 2, .1, .1))
plot_CI(alpha_1, beta_1, conf = 0.05, lambda_true, n=2)
plot_CI(alpha_2, beta_2, conf = 0.05, lambda_true, n=5)
plot_CI(alpha_3, beta_3, conf = 0.05, lambda_true, n=15)
plot_CI(alpha_4, beta_4, conf = 0.05, lambda_true, n=75)
```

## HPD interval
### prior distribution: Gamma(1, 1)
95% confidence interval for the parameter $\lambda$

```{r}
plot_HPDCI <- function(y, lambda, alpha, beta, conf, lambda_true, n) {
    y <- rpois(n, lambda_true)
    posterior <- dgamma(lambda, alpha, beta) * 
    sapply(lambda, function(lambda){prod(dpois(y, lambda))})
    posterior <- posterior / sum(posterior)
    dens <- density(lambda, weights = posterior)
    hpd_region <- hdi(dens, prob = 0.95, allowSplit = TRUE)
    lower <- hpd_region[1, 1]
    upper <- hpd_region[1, 2]
    c(lower, upper)

    plot(lambda, dgamma(lambda, alpha + sum(y), beta + n), 
    type = 'l', lwd = 2, col = 'violet',
    ylim = c(0, 2), xlim = c(0, 10), xlab = expression(lambda),
    ylab = expression(paste('p(', lambda, '|y)')))
    y_post <- dgamma(lambda, alpha + sum(y), beta + n)
    x_coord <- c(lower, lambda[lambda >= lower & lambda <= upper], upper)
    y_coord <- c(0, y_post[lambda >= lower & lambda <= upper], 0)
    polygon(x_coord, y_coord, col = 'pink', lwd = 2,  border = 'violet')
    abline(v = lambda_true, lty = 2)
    text(x = 7, y = 1.5, paste0('n=', n), cex = 1.75)
}
```

```{r}
set.seed(123)
alpha <- 1
beta <- 1
lambda_vals <- seq(0, 15, by = 0.001) 
pdf("/Users/yanjintong/2023/STAT520C/HPD_1.pdf", width=8, height=8)
par(mfrow = c(2, 2), mar = c(2, 2, .1, .1))
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 2)
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 5)
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 15)
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 75)
```

### prior distribution: Gamma(8, 1)
95% confidence interval for the parameter $\lambda$
```{r}
set.seed(123)
alpha <- 8
beta <- 1
lambda_vals <- seq(0, 15, by = 0.001) 
pdf("/Users/yanjintong/2023/STAT520C/HPD_2.pdf", width=8, height=8)
par(mfrow = c(2, 2), mar = c(2, 2, .1, .1))
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 2)
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 5)
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 15)
plot_HPDCI(y, lambda_vals, alpha, beta, 0.05, lambda_true, 75)
```
